---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  message = FALSE,
  warning = FALSE
)
```

# About `meps.hc`

[![Travis-CI Build Status](https://travis-ci.org/jjchern/meps.hc.svg?branch=master)](https://travis-ci.org/jjchern/meps.hc)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/jjchern/meps.hc?branch=master&svg=true)](https://ci.appveyor.com/project/jjchern/meps.hc)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/meps.hc)](https://cran.r-project.org/package=meps.hc)

The goal of `meps.hc` is to wrap the Annual Consolidated Data Files from the 
Medical Expenditure Panel Survey (`meps`) Household Component (`hc`) in an R 
data package. All variable labels and value labels are included.

For more information about the consolidated MEPS-HC files, see [the AHRQ webpages](https://meps.ahrq.gov/mepsweb/data_stats/download_data_files_results.jsp?cboDataYear=All&cboDataTypeY=1%2CHousehold+Full+Year+File&buttonYearandDataType=Search&cboPufNumber=All&SearchTitle=Consolidated+Data), or the example code from the [`HHS-AHRQ/MEPS` repo](https://github.com/HHS-AHRQ/MEPS) by [`@e-mitchell`](https://github.com/e-mitchell).

For another attempt of downloading the MEPS data, see [the `ajdamico/asdfree` repo ](https://github.com/ajdamico/asdfree/tree/master/Medical%20Expenditure%20Panel%20Survey) by [`@ajdamico`](https://github.com/ajdamico).

Currently the package includes data from 2011-2014. The rest of the files (1996-2010) will be packaged soon.

# All MEPS-Related R data packages:

- Full Year Consolidated Files: [`meps.hc`](https://github.com/jjchern/meps.hc)
- Person Round Plan Files: [`meps.prpl`](https://github.com/jjchern/meps.prpl)
- Two-Years Longitudinal Files: [`meps.panel`](https://github.com/jjchern/meps.panel)

# Installation

```R
# install.packages("devtools")
devtools::install_github("jjchern/meps.hc")

# To uninstall the package, use:
# remove.packages("meps.hc")
```

# Usage

```{r}
# Load tibble for better printout
library(tibble)

meps.hc::f2014
```

